Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries

### 2019-05-16 docs/PLoS_manuscript.tex
Updated for appeal process

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Community-based TURF-reserves in small-scale fisheries} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Carlos Villaseñor-Derbez\textsuperscript{1\Yinyang*},
Eréndira Aceves-Bueno\textsuperscript{1,2\Yinyang},
Stuart Fulton\textsuperscript{3\ddag},
Alvin Suarez\textsuperscript{3\ddag},
Arturo Hernández-Velasco\textsuperscript{3\ddag},
Jorge Torre\textsuperscript{3\ddag},
Fiorenza Micheli\textsuperscript{4\ddag}
\\
\bigskip
\textbf{1} Bren School of Environmental Science \& Management, University of California, Santa Barbara, Santa Barbara, CA, USA
\\
\textbf{2} Nicholas School of the Environment, Duke University, Beaufort, NC, USA
\\
\textbf{3} Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico
\\
\textbf{4} Hopkins Marine Station and Stanford Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
\\
\bigskip

@@

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \cite{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce, given the heterogeneity in target species and number of actors involved \cite{costello_2012}. One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves\cite{afflerbach_2014,gelcich_2015,lester_2017}.

@@

Conservation science has shown how well implemented and enforced marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \cite{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \cite{gelcich_2008,gelcich_2012}. The benefits resulting from reserves established within TURFs (\emph{i.e.} TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \cite{gelcich_2015}. Although in theory these systems are expected to be successful \cite{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.

Recent changes in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries. In Mexico, a legal framework created in 2012 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (\emph{Zona de Refugio Pesquero}, described in more detail below; \cite{nom}). Since 2012, 45 old and new marine reserves have gained legal recognition as Fish Refuges. Of these, 18 were originally implemented within TURFs. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico. The objective of this work is twofold. First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries. The second objective is to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves.
%We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{An interdisciplinary evaluation of community-based TURF-reserves} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Carlos Villaseñor-Derbez\textsuperscript{1\Yinyang*},
Eréndira Aceves-Bueno\textsuperscript{1,2\Yinyang},
Stuart Fulton\textsuperscript{3\ddag},
Alvin Suarez\textsuperscript{3\ddag},
Arturo Hernández-Velasco\textsuperscript{3\ddag},
Jorge Torre\textsuperscript{3\ddag},
Fiorenza Micheli\textsuperscript{4\ddag}
\\
\bigskip
\textbf{1} Bren School of Environmental Science \& Management, University of California, Santa Barbara, Santa Barbara, CA, USA
\\
\textbf{2} Nicholas School of the Environment, Duke University, Beaufort, NC, USA
\\
\textbf{3} Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico
\\
\textbf{4} Hopkins Marine Station and Stanford Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
\\
\bigskip

@@

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \cite{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce, given the large number of participants, heterogeneity in fleet, gears, and targeted species; as well as seasonality and geographical distribution \cite{costello_2012,jacquet_2008}. One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves \cite{afflerbach_2014,gelcich_2015,lester_2017}.

@@

Conservation science has shown how well implemented and enforced marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \cite{halpern_2002,lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \cite{gelcich_2008,gelcich_2012}. The benefits resulting from reserves established within TURFs (\emph{i.e.} TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \cite{gelcich_2015}. Although in theory these systems are expected to be successful \cite{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.

Recent changes in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries. In Mexico, a legal framework created in 2012 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (\emph{Zona de Refugio Pesquero}, described in more detail below; \cite{nom}). Since 2012, 45 old and new marine reserves have gained legal recognition as Fish Refuges. Of these, 18 were originally implemented within TURFs. However, their effectiveness has not yet been reported in the scientific literature.

Marine reserves have largely been evaluated from an ecological perspective, focusing on ecological indicators such as biomass and species richness or spillover\cite{halpern_2002,lester_2009,krueck_2017-J1}. However, TURF-reserves are intricate social-ecological systems, oftem implemented with a combination of ecological and social objectives. A customary evaluation of TURF-reserves is unlikely to provide an accurate representation of the changes, benefits, and limitations of such an intervention.

The objective of this work is twofold. First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries. The second objective is to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves. We combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico.  Our work highlights the benefits of taking an interdisciplinary approach to marine reserve evaluation, and provides a robust evaluation of TURF-reserve effectiveness that may inform adaptive management and future interventions.
